### 2022-03-05 paper/output/document/thesis_0_abstract.tex
added will's fixes to ch5

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}abstract}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
AI agents are beginning to make decisions without human supervision in increasingly consequential 
contexts like healthcare, policing, and driving. These decisions are inevitably ethically tinged, 
but most AI agents navigating the world today are not explicitly guided by ethics.
Warnings from regulators, philosophers, and computer scientists about the 
dangers of unethical artificial intelligence, from science-fiction killer robots to criminal
sentencing algorithms prejudiced against people of color, have spurred interest in automated ethics, or the development 
of machines that can perform ethical reasoning. Much prior work in automated ethics approaches the 
problem from a computational perspective and rarely engages with philosophical literature on ethics, despite
its relevance to the development of AI agents that can
responsibly navigate the world. If automated ethics draws on sophisticated philosophical literature, the ethical reasoning
underlying such decisions will be more nuanced, precise, consistent, and trustworthy. However, faithfully translating complex ethical theories
from natural language to the rigid syntax of a computer program poses technical and philosophical 
challenges.

In this thesis, I present an implementation of automated Kantian
ethics that is faithful to the Kantian philosophical tradition. Of the three major ethical
traditions, Kant's categorical imperative is the most natural to formalize because it is an inviolable 
formal rule that requires less situational awareness than other ethical theories. I formalize Kant's categorical imperative 
in Carmo and Jones's Dyadic Deontic Logic, implement this formalization 
in the Isabelle/HOL theorem prover, and develop a testing framework to evaluate how well 
my implementation coheres with expected properties of Kantian ethics, as established in the literature. 
I also use my system to reason about two ethical dilemmas used to criticize Kantian ethics: the difference
between lying and joking and the example of a murderer knocking on your door asking about the location of their
intended victim. Finally, I examine the philosophical implications of this system, exploring its limitations 
and its potential to help both AI agents and human beings better reason about ethics. This work serves 
as an early proof-of-concept for philosophically mature AI agents and is one step towards the development 
of responsible, trustworthy artificial intelligence.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_0_abstract.thy%:%
%:%19=6%:%
%:%20=7%:%
%:%21=8%:%
%:%22=9%:%
%:%23=10%:%
%:%24=11%:%
%:%25=12%:%
%:%26=13%:%
%:%27=14%:%
%:%28=15%:%
%:%29=16%:%
%:%30=17%:%
%:%31=18%:%
%:%32=19%:%
%:%33=20%:%
%:%34=21%:%
%:%35=22%:%
%:%36=23%:%
%:%37=24%:%
%:%38=25%:%
%:%39=26%:%
%:%40=27%:%
%:%41=28%:%
%:%42=29%:%
%:%43=30%:%
%:%44=31%:%
%:%45=32%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}abstract}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\begin{isamarkuptext}%
AI agents are beginning to make decisions without human supervision in increasingly consequential 
contexts like healthcare, policing, and driving. These decisions are inevitably ethically tinged, 
but most AI agents navigating the world today are not explicitly guided by ethics.
Regulators, philosophers, and computer scientists are raising the alarm about the 
dangers of unethical artificial intelligence, from science-fiction killer robots to criminal
sentencing algorithms prejudiced against people of color. This is spurring interest in automated ethics, or the development 
of machines that can perform ethical reasoning. Much prior work in automated ethics approaches the 
problem from a computational perspective and rarely engages with philosophical literature on ethics, despite
its relevance to the development of AI agents that can
responsibly navigate the world. If automated ethics draws on sophisticated philosophical literature, the ethical reasoning
underlying such decisions will be more nuanced, precise, consistent, and trustworthy. However, faithfully translating complex ethical theories
from natural language to the rigid syntax of a computer program poses technical and philosophical 
challenges.

In this thesis, I present an implementation of automated Kantian
ethics that is faithful to the Kantian philosophical tradition. Given an appropriately represented 
action and minimal factual background, my system can judge the action as morally obligatory, permissible, or prohibited.
To accomplish this, I formalize Kant's categorical imperative, or moral rule,
in Carmo and Jones's Dyadic Deontic Logic, implement this formalization 
in the Isabelle/HOL theorem prover, and develop a testing framework to evaluate how well 
my implementation coheres with expected properties of Kantian ethics, as established in the literature. 
I also use my system to derive philosophically sophisticated and nuanced solutions to two central ethical 
dilemmas in Kantian ethical literature: the difference
between lying and joking and the example of a murderer knocking on your door asking about the location of their
intended victim. Finally, I examine the philosophical implications of my system, exploring its limitations 
and its potential to help AI agents, academic philosophers, and ordinary human beings better reason about ethics. This work serves 
as an early proof-of-concept for philosophically mature AI agents and is one step towards the development 
of responsible, trustworthy artificial intelligence.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_0_abstract.thy%:%
%:%19=6%:%
%:%20=7%:%
%:%21=8%:%
%:%22=9%:%
%:%23=10%:%
%:%24=11%:%
%:%25=12%:%
%:%26=13%:%
%:%27=14%:%
%:%28=15%:%
%:%29=16%:%
%:%30=17%:%
%:%31=18%:%
%:%32=19%:%
%:%33=20%:%
%:%34=21%:%
%:%35=22%:%
%:%36=23%:%
%:%37=24%:%
%:%38=25%:%
%:%39=26%:%
%:%40=27%:%
%:%41=28%:%
%:%42=29%:%
%:%43=30%:%
%:%44=31%:%
%:%45=32%:%
%:%46=33%:%

### 2022-03-05 paper/output/document/thesis_1_intro.tex
added will's fixes to ch5

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}intro}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Introduction\label{intro}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As AI agents become more sophisticated and less dependent on humans, interest begins to mount
in the development of computers that can perform ethical reasoning, also known as automated moral agents. 
AI agents are making decisions in increasingly 
consequential contexts, such as healthcare, driving, and criminal sentencing, and therefore 
must perform ethical reasoning in order to navigate moral dilemmas. For example, self-driving
cars may face less extreme versions of the following moral dilemma: an autonomous vehicle approaching 
an intersection fails to notice pedestrians in the crosswalk until it is too late to brake. The car 
can either continue on its course, running over and killing three pedestrians, or it can swerve to 
hit the car in the next lane, killing the single passenger inside it. While this example is (hopefully) 
not typical of the operation of a self-driving car, every decision that such an AI agent makes, from 
avoiding congested freeways to carpooling, is morally tinged. Not only do AI agents routinely make decisions with 
ethical implications without explicitly performing ethical
reasoning, in many cases, they do so without human supervision. For example, the Alleghany Family Screening 
tool can automatically trigger an investigation into a potential case of child neglect, a decision that 
can uproot entire families and is known to be biased against poor people of color \citep{eubanks}. 
This motivates the need for machine ethics (also called automated ethics), 
or the study of how to develop machines that can perform robust, sophisticated ethical reasoning.

@@

A lack of engagement with prior philosophical literature also makes automated moral agents less 
explainable, or interpretable by human observers. One example of this is Delphi, a language model that uses deep 
learning to make moral judgements based on a training dataset of ethical decisions made by humans \citep{delphi}. 
Early versions of Delphi gave unexpected results, such as declaring that the user should commit 
genocide if it makes everyone happy \citep{verge}. Moreover, because no explicit ethical theory underpins 
Delphi's judgements, human beings cannot analytically determine why Delphi thinks genocide is obligatory
or where its reasoning may have gone wrong. 
Machine learning approaches like Delphi often cannot explain their decisions to a human being and, 
in the extreme case, are black box algorithms. This reduces human trust in a machine's controversial 
ethical judgements. If a machine prescribes killing one person to save three without justifying this 
decision, it is difficult to trust this judgement enough to act on it or endorse a machine acting on it. 
The high stakes of automated ethics require explainability to build trust and catch mistakes.

While automated ethics should draw on philosophical literature, in practice, automating an ethical 
theory is a technical and philosophical challenge. Intuitive computational approaches explored 
previously, such as representing ethics as a constraint satisfaction problem \citep{csp} or reinforcement 
learning algorithm \citep{util1}, fail to capture philosophically plausible ethical theories. For 
example, encoding ethics as a Markov Decision Process assumes that ethical reward can be aggregated 
according to some discounted sum, but many philosophers reject this notion of aggregation \citep{consequentialismsep}. 
Approaches that begin with an ethical theory, instead of a computational tool, must contend with the 
fact that ethical theories are almost always described in natural language and must be made
precise enough to represent to a computer. Even once ethics is translated from natural 
language to program syntax, the factual background given to the machine, such as the description of 
an ethical dilemma, is equally as important in determining the machine's decisions. Another complication
is that philosophers do not agree that on a single choice of ethical theory. Philosophers who
so agree that a specific ethical theory, like Kantian ethics, is true, still debate the theory's 
details.\footnote{For examples of these debates in the case of Kantian ethics, see Section Joking
and Section Murderer.} Moreover, even once reasoning within a 
particular ethical theory is automated, those who disagree with that theory will disagree with the 
system's judgements.

This thesis presents a proof-of-concept implementation of philosophically faithful automated Kantian ethics. 
I formalize Kant's categorical imperative, or moral rule, as an axiom 
in Carmo and Jones' Dyadic Deontic Logic (DDL), a modal logic designed to reason about 
obligation \citep{CJDDL}. I implement my formalization in Isabelle/HOL, an interactive theorem prover 
that can automatically verify and generate proofs in user-defined logics \citep{isabelle}. Finally, 
I use Isabelle to automatically prove theorems (such as, ``murder is wrong'') in my new logic, 
generating results derived from the categorical imperative. Because my system automates reasoning in 
a logic that represents Kantian ethics, it automates Kantian ethical reasoning. Once equipped with 
minimal factual background, it can classify actions as prohibited, permissible or obligatory. I 
make the following contributions:%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\begin{enumerate}
\item In Section \ref{whykant}, I make a philosophical argument for why Kantian ethics is the most natural of the three major
ethical traditions (deontology, virtue ethics, utilitarianism) to formalize.

@@

\item In Section \ref{computationalethics}, I present new ethical insights discovered using my system and argue that
computational methods like the one presented in this paper can help philosophers address ethical problems.
Not only can my system help machines reason about ethics, but it can also help philosophers make philosophical
progress.
\end{enumerate}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
I choose to formalize Kant's moral rule in Carmo and Jones' Dyadic Deontic Logic (DDL) \citep{CJDDL}. Deontic 
logic is a modal logic that can express obligation, or morally binding requirements. Traditional modal 
logics include the necessitation operator, denoted as $\Box$. In modal logic using the Kripke semantics, 
$\Box p$ is true at world $w$ if $p$ is true at all worlds that neighbor $w$ \citep{cresswell}. Modal 
logics  also contain the possibility operator $\diamond$, where $\diamond \, p \iff \neg (\Box (\neg p))$ 
and operators of propositional logic like $\neg, \wedge, \vee, \rightarrow$. I use DDL, in which
the dyadic obligation operator $O\{A \vert B\}$ represents the sentence ``A is obligated in the context B.'' 
The introduction of context allows DDL to express more nuanced reasoning. DDL is both deontic and modal, 
so sentences like $O\{A \vert B\}$ are terms that can be true or false at a world. For example, the 
sentence $O \{ \text{steal} \vert \text{when rich}\}$ is true at a world if stealing when rich is 
obligated at that particular world.

I automate Kantian ethics because it is the most natural to formalize, as I argue in Section WhyKant. 
Kant presents three versions of a single moral rule, known as the categorical imperative, from which 
all moral judgements can be derived. I implement a version of this rule called the Formula of Universal 
Law (FUL), which states that people should only act on those principles that can be acted on by all 
people without contradiction. For example, in a world where everyone falsely promises to repay a loan, 
lenders will no longer believe these promises and will stop offering loans. Therefore, not everyone 
can simultaneously falsely promise to repay a loan, so the FUL thus prohibits this act.

Prior work by Benzmüller, Farjami, and Parent \citep{logikey, BFP} implements DDL in Isabelle/HOL and 
I add the Formula of Universal Law as an axiom on top of their library. The resulting Isabelle theory 
can automatically or semi-automatically generate proofs in a new logic that has the categorical 
imperative as an axiom. Because proofs in this logic are derived from the categorical imperative, 
they judge actions as obligated, prohibited, or permissible. Moreover, because interactive 
theorem provers are designed to be interpretable, my system is explainable. Isabelle can list 
the axioms and facts it used to generate an ethical judgement, and, in some cases, construct 
human-readable proofs. In Sections Joking and Murderer, I use my system to arrive at 
sophisticated solutions to two ethical dilemmas often used in critiques of Kantian ethics. Because 
my system is faithful to philosophical literature, it is able to provide nuanced answers to these paradoxes
that require a deep understanding of Kantian ethics.

In addition to presenting the above logic and implementation, I also contribute a testing framework 
that evaluates how well my formalization coheres with philosophical literature. I formalize expected 
properties of Kantian ethics as sentences in my logic, such as the property that obligations cannot 
contradict each other. I represent each of these properties as a sentence in my logic that my system 
should be able to prove or refute. I run the tests by using Isabelle to automatically find proofs or 
countermodels for the test statements. For example, my implementation passes the contradictory 
obligations test because it is able to prove the sentence $\neg (O\{A|B\} \wedge O\{\neg A | B\})$. 
I find that my system outperforms the control group of raw DDL, without any moral axioms added, and 
Moshe Kroy's prior attempt at formalizing Kantian ethics in deontic logic \citep{kroy}.

@@

A machine that can evaluate the moral status of a maxim can not only help machines better reason about ethics, 
but it can also help philosophers 
better study philosophy. I argue for ``computational ethics," or the use of computational tools to 
make philosophical progress. I demonstrate the potential of computational ethics by presenting a 
philosophical insight about which kinds of maxims are appropriate for ethical consideration that I 
discovered using my system. The process of building and interacting with a computer that can reason 
about ethics helped me, a human philosopher, arrive at a philosophical conclusion that has implications for practical
reason and philosophy of doubt. Thus, my system can be used in two distinct ways. First, my system can help
automated agents navigate the world, which I will refer to as automated ethics or machine ethics interchangeably. Second, 
my system help human philosophers reason about philosophy, which I call computational ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_1_intro.thy%:%
%:%24=6%:%
%:%36=8%:%
%:%37=9%:%
%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%42=14%:%
%:%43=15%:%
%:%44=16%:%
%:%45=17%:%
%:%46=18%:%
%:%47=19%:%
%:%48=20%:%
%:%49=21%:%
%:%50=22%:%
%:%51=23%:%
%:%52=24%:%
%:%53=25%:%
%:%54=26%:%
%:%55=27%:%
%:%56=28%:%
%:%57=29%:%
%:%58=30%:%
%:%59=31%:%
%:%60=32%:%
%:%61=33%:%
%:%62=34%:%
%:%63=35%:%
%:%64=36%:%
%:%65=37%:%
%:%66=38%:%
%:%67=39%:%
%:%68=40%:%
%:%69=41%:%
%:%70=42%:%
%:%71=43%:%
%:%72=44%:%
%:%73=45%:%
%:%74=46%:%
%:%75=47%:%
%:%76=48%:%
%:%77=49%:%
%:%78=50%:%
%:%79=51%:%
%:%80=52%:%
%:%81=53%:%
%:%82=54%:%
%:%83=55%:%
%:%84=56%:%
%:%85=57%:%
%:%86=58%:%
%:%87=59%:%
%:%88=60%:%
%:%89=61%:%
%:%90=62%:%
%:%91=63%:%
%:%92=64%:%
%:%93=65%:%
%:%94=66%:%
%:%95=67%:%
%:%96=68%:%
%:%97=69%:%
%:%98=70%:%
%:%99=71%:%
%:%100=72%:%
%:%101=73%:%
%:%102=74%:%
%:%103=75%:%
%:%104=76%:%
%:%105=77%:%
%:%106=78%:%
%:%107=79%:%
%:%108=80%:%
%:%109=81%:%
%:%113=84%:%
%:%114=85%:%
%:%115=86%:%
%:%116=87%:%
%:%117=88%:%
%:%118=89%:%
%:%119=90%:%
%:%120=91%:%
%:%121=92%:%
%:%122=93%:%
%:%123=94%:%
%:%124=95%:%
%:%125=96%:%
%:%126=97%:%
%:%127=98%:%
%:%128=99%:%
%:%129=100%:%
%:%130=101%:%
%:%131=102%:%
%:%132=103%:%
%:%133=104%:%
%:%134=105%:%
%:%135=106%:%
%:%136=107%:%
%:%137=108%:%
%:%141=112%:%
%:%142=113%:%
%:%143=114%:%
%:%144=115%:%
%:%145=116%:%
%:%146=117%:%
%:%147=118%:%
%:%148=119%:%
%:%149=120%:%
%:%150=121%:%
%:%151=122%:%
%:%152=123%:%
%:%153=124%:%
%:%154=125%:%
%:%155=126%:%
%:%156=127%:%
%:%157=128%:%
%:%158=129%:%
%:%159=130%:%
%:%160=131%:%
%:%161=132%:%
%:%162=133%:%
%:%163=134%:%
%:%164=135%:%
%:%165=136%:%
%:%166=137%:%
%:%167=138%:%
%:%168=139%:%
%:%169=140%:%
%:%170=141%:%
%:%171=142%:%
%:%172=143%:%
%:%173=144%:%
%:%174=145%:%
%:%175=146%:%
%:%176=147%:%
%:%177=148%:%
%:%178=149%:%
%:%179=150%:%
%:%180=151%:%
%:%181=152%:%
%:%182=153%:%
%:%183=154%:%
%:%184=155%:%
%:%185=156%:%
%:%186=157%:%
%:%187=158%:%
%:%188=159%:%
%:%189=160%:%
%:%190=161%:%
%:%191=162%:%
%:%192=163%:%
%:%193=164%:%
%:%194=165%:%
%:%195=166%:%
%:%196=167%:%
%:%197=168%:%
%:%198=169%:%
%:%199=170%:%
%:%200=171%:%
%:%201=172%:%
%:%202=173%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}intro}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Introduction\label{intro}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As AI agents become more sophisticated and less dependent on humans, interest begins to mount
in the development of computers that can perform ethical reasoning, also known as automated moral agents. 
AI agents are making decisions in increasingly 
consequential contexts, such as healthcare, driving, and criminal sentencing, and therefore 
must perform ethical reasoning in order to navigate moral dilemmas. For example, self-driving
cars may face less extreme versions of the following moral dilemma: an autonomous vehicle approaching 
an intersection fails to notice pedestrians in the crosswalk until it is too late to brake. The car 
can either continue on its course, running over and killing three pedestrians, or it can swerve to 
hit the car in the next lane, killing the single passenger inside it. While this example is (hopefully) 
not typical of the operation of a self-driving car, every decision that such an AI agent makes, from 
avoiding congested freeways to carpooling, is morally tinged. Not only do AI agents routinely make decisions with 
ethical implications without explicitly performing ethical
reasoning, in many cases they do so without human supervision. For example, the Alleghany Family Screening 
tool can automatically trigger an investigation into a potential case of child neglect, a decision that 
can uproot entire families and is known to be biased against poor people of color \citep{eubanks}. 
This motivates the need for machine ethics (also called automated ethics), 
or the study of how to develop machines that can perform robust, sophisticated ethical reasoning.

@@

A lack of engagement with prior philosophical literature also makes automated moral agents less 
explainable, or interpretable by human observers. One example of this is Delphi, an implementation of automated ethics that uses deep 
learning to make moral judgements based on a training dataset of ethical decisions made by humans \citep{delphi}. 
Early versions of Delphi gave unexpected results, such as declaring that the user should commit 
genocide if it makes everyone happy \citep{verge}. Moreover, because no explicit ethical theory underpins 
Delphi's judgements, human beings cannot analytically determine why Delphi thinks genocide is obligatory
or where its reasoning may have gone wrong. 
Machine learning approaches like Delphi often cannot explain their decisions to a human being and, 
in the extreme case, are black box algorithms. This reduces human trust in a machine's controversial 
ethical judgements. If a machine prescribes killing one person to save three without justifying this 
decision, it is difficult to trust this judgement. 
The high stakes of automated ethics require explainability to build trust and catch mistakes.

While automated ethics should draw on philosophical literature, in practice, automating an ethical 
theory is a technical and philosophical challenge. Intuitive computational approaches explored 
previously, such as representing ethics as a constraint satisfaction problem \citep{csp} or reinforcement 
learning algorithm \citep{util1}, fail to capture philosophically plausible ethical theories. For 
example, encoding ethics as a Markov Decision Process assumes that ethical reward can be aggregated 
according to some discounted sum\footnote{Markov Decision Processes usually assume that the total
reward of a system is the discounted sum of the reward at each state, given by $r_i$. Formally, total
reward $R=\sum_0^{\infty}\gamma^ir_i$.}, but many philosophers reject this notion of aggregation \citep{consequentialismsep}. 
On the other hand, approaches that begin with an ethical theory, instead of a computational method, must contend with the 
fact that ethical theories are almost always described in natural language and must be made
precise enough to represent to a computer. Even once ethics is translated from natural 
language to program syntax, the factual background given to the machine, such as the description of 
an ethical dilemma, plays a great role in the machine's decisions. Another complication
is that philosophers do not agree that on a single choice of ethical theory. Even philosophers who
agree that a specific ethical theory, like Kantian ethics, is true, still debate the theory's 
details.\footnote{For examples of these debates in the case of Kantian ethics, see Section Joking
and Section Murderer.} Moreover, even once reasoning within a 
particular ethical theory is automated, those who disagree with that theory will disagree with the 
system's judgements.

This thesis presents a proof-of-concept implementation of philosophically faithful automated Kantian ethics. 
I formalize Kant's categorical imperative, or moral rule, as an axiom 
in Dyadic Deontic Logic (DDL), a modal logic designed to reason about 
obligation \citep{CJDDL}. I implement my formalization in Isabelle/HOL, an interactive theorem prover 
that can automatically verify and generate proofs in user-defined logics \citep{isabelle}. Finally, 
I use Isabelle to automatically prove theorems (such as, ``murder is wrong'') in my new logic, 
generating results derived from the categorical imperative. Because my system automates reasoning in 
a logic that represents Kantian ethics, it automates Kantian ethical reasoning. Once equipped with 
minimal factual background, it can classify actions as prohibited, permissible or obligatory. I 
make the following contributions:%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\begin{enumerate}
\item In Section \ref{whykant}, I make a philosophical argument for why Kantian ethics is the most natural of the three major
ethical traditions (deontology, virtue ethics, utilitarianism) to formalize.

@@

\item In Section \ref{computationalethics}, I present new ethical insights discovered using my system and argue that
computational methods like the one presented in this paper can help philosophers address ethical problems.
Not only can my system help machines reason about ethics, but it can also help philosophers make philosophical
progress.
\end{enumerate}%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
Deontic logic is a modal logic that can express obligation, or morally binding requirements. Traditional modal 
logics include the necessitation operator, denoted as $\Box$. In modal logic using the Kripke semantics, 
$\Box p$ is true at world $w$ if $p$ is true at all worlds that neighbor $w$ \citep{cresswell}. Modal 
logics  also contain the possibility operator $\diamond$, where $\diamond \, p \longleftrightarrow \neg (\Box (\neg p))$ 
and operators of propositional logic like $\neg, \wedge, \vee, \rightarrow$. I use DDL, in which
the dyadic obligation operator $O\{A \vert B\}$ represents the sentence ``A is obligated in the context B.'' 
The introduction of context allows DDL to express more nuanced reasoning. DDL is both deontic and modal, 
so sentences like $O\{A \vert B\}$ are terms that can be true or false at a world. For example, the 
sentence $O \{ \text{steal} \vert \text{when rich}\}$ is true at a world if stealing when rich is 
obligated at that particular world.

I automate Kantian ethics because it is the most natural to formalize, as I argue in Section \ref{whykant}. 
Kant presents three versions of a single moral rule, known as the categorical imperative, from which 
all moral judgements can be derived. I implement a version of this rule called the Formula of Universal 
Law (FUL), which states that people should only act on those principles that can be acted on by all 
people without contradiction. For example, in a world where everyone falsely promises to repay a loan, 
lenders will no longer believe these promises and will stop offering loans. Therefore, not everyone 
can simultaneously falsely promise to repay a loan, so the FUL thus prohibits this act.

Prior work by Benzmüller, Farjami, and Parent \citep{logikey, BFP} implements DDL in Isabelle/HOL and 
I add the Formula of Universal Law as an axiom on top of their library. The resulting Isabelle theory 
can automatically or semi-automatically generate proofs in a new logic that has the categorical 
imperative as an axiom. Because proofs in this logic are derived from the categorical imperative, 
they judge actions as obligated, prohibited, or permissible. Moreover, because interactive 
theorem provers are designed to be interpretable, my system is explainable. Isabelle can list 
the axioms and facts it used to generate an ethical judgement, and, in some cases, construct 
human-readable proofs. In Chapter \ref{applications}, I use my system to derive
sophisticated solutions to two ethical dilemmas often used in critiques of Kantian ethics. Because 
my system is faithful to philosophical literature, it is able to provide nuanced answers to these paradoxes
that require a deep understanding of Kantian ethics.

In addition to presenting the above logic and implementation, I also contribute a testing framework 
that evaluates how well my formalization coheres with philosophical literature. I formalize expected 
properties of Kantian ethics as sentences in my logic, such as the property that obligations cannot 
contradict each other. I represent each of these properties as a sentence in my logic that my system 
should be able to prove or refute. To run the tests, I use Isabelle to automatically find proofs or 
countermodels for the test statements. For example, my implementation passes the contradictory 
obligations test because it is able to prove the sentence $\neg (O\{A|B\} \wedge O\{\neg A | B\})$. 
I find that my system outperforms the control group of raw DDL, without any moral axioms added, and 
Moshe Kroy's prior attempt at formalizing Kantian ethics in deontic logic \citep{kroy}.

@@

A machine that can evaluate the moral status of a maxim can not only help machines better reason about ethics, 
but it can also help philosophers better study philosophy. I argue for ``computational ethics,'' or the use of computational tools to 
make philosophical progress, analogous to computational drug discovery or neuroscience. 
I demonstrate the potential of computational ethics by presenting a 
philosophical insight about which kinds of maxims are appropriate for ethical consideration that I 
discovered using my system. The process of building and interacting with a computer that can reason 
about ethics helped me, a human philosopher, arrive at a philosophical conclusion that has implications for practical
reason and philosophy of doubt. Thus, my system can be used in two distinct ways. First, my system can help
automated agents navigate the world, which I will refer to as automated ethics or machine ethics interchangeably. Second, 
my system help human philosophers reason about philosophy, which I call computational ethics. In Section 
\ref{ordinarypeople}, I consider how computational ethics can help not only professional philosophers,
but also augment the everyday ethical reasoning that ordinary people perform as we navigate the world.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\end{isabellebody}%
\endinput
%:%file=~/Desktop/cs91r/paper/thesis_1_intro.thy%:%
%:%24=6%:%
%:%36=8%:%
%:%37=9%:%
%:%38=10%:%
%:%39=11%:%
%:%40=12%:%
%:%41=13%:%
%:%42=14%:%
%:%43=15%:%
%:%44=16%:%
%:%45=17%:%
%:%46=18%:%
%:%47=19%:%
%:%48=20%:%
%:%49=21%:%
%:%50=22%:%
%:%51=23%:%
%:%52=24%:%
%:%53=25%:%
%:%54=26%:%
%:%55=27%:%
%:%56=28%:%
%:%57=29%:%
%:%58=30%:%
%:%59=31%:%
%:%60=32%:%
%:%61=33%:%
%:%62=34%:%
%:%63=35%:%
%:%64=36%:%
%:%65=37%:%
%:%66=38%:%
%:%67=39%:%
%:%68=40%:%
%:%69=41%:%
%:%70=42%:%
%:%71=43%:%
%:%72=44%:%
%:%73=45%:%
%:%74=46%:%
%:%75=47%:%
%:%76=48%:%
%:%77=49%:%
%:%78=50%:%
%:%79=51%:%
%:%80=52%:%
%:%81=53%:%
%:%82=54%:%
%:%83=55%:%
%:%84=56%:%
%:%85=57%:%
%:%86=58%:%
%:%87=59%:%
%:%88=60%:%
%:%89=61%:%
%:%90=62%:%
%:%91=63%:%
%:%92=64%:%
%:%93=65%:%
%:%94=66%:%
%:%95=67%:%
%:%96=68%:%
%:%97=69%:%
%:%98=70%:%
%:%99=71%:%
%:%100=72%:%
%:%101=73%:%
%:%102=74%:%
%:%103=75%:%
%:%104=76%:%
%:%105=77%:%
%:%106=78%:%
%:%107=79%:%
%:%108=80%:%
%:%109=81%:%
%:%110=82%:%
%:%111=83%:%
%:%115=86%:%
%:%116=87%:%
%:%117=88%:%
%:%118=89%:%
%:%119=90%:%
%:%120=91%:%
%:%121=92%:%
%:%122=93%:%
%:%123=94%:%
%:%124=95%:%
%:%125=96%:%
%:%126=97%:%
%:%127=98%:%
%:%128=99%:%
%:%129=100%:%
%:%130=101%:%
%:%131=102%:%
%:%132=103%:%
%:%133=104%:%
%:%134=105%:%
%:%135=106%:%
%:%136=107%:%
%:%137=108%:%
%:%138=109%:%
%:%139=110%:%
%:%143=114%:%
%:%144=115%:%
%:%145=116%:%
%:%146=117%:%
%:%147=118%:%
%:%148=119%:%
%:%149=120%:%
%:%150=121%:%
%:%151=122%:%
%:%152=123%:%
%:%153=124%:%
%:%154=125%:%
%:%155=126%:%
%:%156=127%:%
%:%157=128%:%
%:%158=129%:%
%:%159=130%:%
%:%160=131%:%
%:%161=132%:%
%:%162=133%:%
%:%163=134%:%
%:%164=135%:%
%:%165=136%:%
%:%166=137%:%
%:%167=138%:%
%:%168=139%:%
%:%169=140%:%
%:%170=141%:%
%:%171=142%:%
%:%172=143%:%
%:%173=144%:%
%:%174=145%:%
%:%175=146%:%
%:%176=147%:%
%:%177=148%:%
%:%178=149%:%
%:%179=150%:%
%:%180=151%:%
%:%181=152%:%
%:%182=153%:%
%:%183=154%:%
%:%184=155%:%
%:%185=156%:%
%:%186=157%:%
%:%187=158%:%
%:%188=159%:%
%:%189=160%:%
%:%190=161%:%
%:%191=162%:%
%:%192=163%:%
%:%193=164%:%
%:%194=165%:%
%:%195=166%:%
%:%196=167%:%
%:%197=168%:%
%:%198=169%:%
%:%199=170%:%
%:%200=171%:%
%:%201=172%:%
%:%202=173%:%
%:%203=174%:%
%:%204=175%:%
%:%205=176%:%
